# (PART) Getting Started {-}

\setcounter{chapter}{0}

# The Basics {#basics}

Before you get too far into the Style Guide, we would like for you to take a moment and ensure that you have the following tools at your disposal.

## Getting Started Checklist {#checklist}
1. Ensure that you have all of necessary accounts and if not, put in a request to Bob ([rpc5102](mailto:rpc5102@psu.edu)).  
    a. [DataCamp](https://www.datacamp.com/)  
    b. [GitHub](https://github.com/)  
    c. [EducationShinyAppTeam](https://github.com/EducationShinyAppTeam) 
    d. [BOAST](https://teams.microsoft.com/l/team/19%3a8423bc25992a4451952d8312b497324d%40thread.skype/conversations?groupId=a83c9d1b-6bcc-46f4-aa25-82b2fad862ec&tenantId=7cf48d45-3ddb-4389-a9c1-c115526eb52e) in Microsoft Teams (tied to your PSU AccessID)
2. Ensure that you have all of the required software  
    a. [`R` (version 3.5.* minimum, version 4.0.0 preferred)](https://cloud.r-project.org/) 
    b. [RStudio Desktop (most current version preferred)](https://rstudio.com/products/rstudio/download/#download)
3. Additional Software that we strongly recommend 
    a. [GitHub Desktop](https://desktop.github.com/)
4. `R` Packages--here are some basic packages that everyone will need; be sure to install their dependencies

    ```{r packageInstall, echo = TRUE, eval = FALSE}
    install.packages(c("devtools", "DT", "ggplot2", "shiny", "shinyBS", 
                       "shinyWidgets", "tidyverse", "tippy"))
    ```

5. A copy of the Sample App ([see below](#sample-app))

## `boastUtils` Package
Bob created the [boastUtils](https://github.com/EducationShinyAppTeam/boastUtils) package to automate much of the design and development process. This will not only reduce the amount of work you'll need to do, it'll also make apps more consistent. 
 
__Starting Summer 2020, you will be required to you make use of this tool.__

Please check out the [package's GitHub page](https://github.com/EducationShinyAppTeam/boastUtils) for instructions on installing and usage.

```{r utilInstall, echo = TRUE, eval = FALSE}
library(devtools)
devtools::install_github("EducationShinyAppTeam/boastUtils")
```

## Sample App  
Bob and Neil have created a Sample App repository that you can use as a template for your own apps. To get started, clone the [Sample\_App](https://github.com/EducationShinyAppTeam/Sample_App) template repo found on GitHub. This will provide you with a skeleton for organizing your files as well as your code.

There are several methods you can use:

- [Direct Download (Basic)](#direct-download)
- [GitHub Desktop (Recommended)](#github-desktop)
- [Command Line (Advanced)](#command-line)
- [RStudio Project](#rstudio-project)

### Direct Download
You can download this repository directly by visiting:

> <https://github.com/EducationShinyAppTeam/Sample_App/archive/master.zip>

### GitHub Desktop
If you are using [GitHub Desktop](https://desktop.github.com/) and have [linked your account](https://help.github.com/en/desktop/getting-started-with-github-desktop/authenticating-to-github) that has access to the [EducationShinyAppTeam](https://github.com/EducationShinyAppTeam) repository, you can do the following from inside GitHub Desktop:  

1. Bring up the Clone Repository Menu (File -> Clone Repository...)  
2. Enter Sample\_App in the search bar and select the option that says Sample\_App (not sampleapp)
3. Click the Choose... button for the local path (this is where you want to the clone to live on your computer)
4. Click the Clone button.

**See also:**

- [GitHub Desktop Help](https://help.github.com/en/desktop)


### Command Line
Enter the following command in your terminal:

```{bash, eval = FALSE}
git clone git@github.com:EducationShinyAppTeam/Sample_App.git
```

**See also:**

- [Git Handbook](https://guides.github.com/introduction/git-handbook/)
- [Resources to learn Git](https://try.github.io/)
- [Happy Git and GitHub for the userR](https://happygitwithr.com/)

### RStudio Project {#rstudio-project}
You can also use RStudio to connect to and copy repositories from GitHub. There are a few extra steps that you'll need to go through compared to other methods. We recommend that you use GitHub Desktop.

1. In your browser, navigate to the repository you want to copy; in this case the [Sample\_App repo](https://github.com/EducationShinyAppTeam/Sample_App).
2. Launch RStudio and create a new project.
3. Select the Version Control option and then Git.
4. Switch back to your browser and click on the Clone or Download button. There should be a line of text starting with `git@github` that you can copy. Do so.
5. Returning to RStudio, paste the get you just copied into the Repository URL field.
6. Adjust the directory name and the subdirectory (where this project will live on your machine) as you see fit.
7. Click the Create Project button.

This method will ensure that you will be able to successfully push and pull changes to GitHub. For additional help, Neil has created a video "[Using RStudio to Connect to GitHub](https://youtu.be/IK0SiEz5vjs)" which walks you through the process.
