\setcounter{chapter}{0}

# Getting Started {#getStarted}

Before you get too far into the Style Guide, we would like for you to take a moment and ensure that you have the following tools at your disposal.

## Getting Started Checklist {#checklist}
1. Ensure that you have all of necessary accounts and if not, put in a request to Bob ([rpc5102](mailto:rpc5102@psu.edu)).  
    a. [DataCamp](https://www.datacamp.com/)  
    b. [GitHub](https://github.com/)  
    c. [EducationShinyAppTeam](https://github.com/EducationShinyAppTeam) 
    d. ~~RStudio Server~~  
          + Upon investigation the TLT and the Eberly RStudio Servers will not be useful for us to do testing. An alternate testing process will be explored.  
    e. [BOAST](https://teams.microsoft.com/l/team/19%3a8423bc25992a4451952d8312b497324d%40thread.skype/conversations?groupId=a83c9d1b-6bcc-46f4-aa25-82b2fad862ec&tenantId=7cf48d45-3ddb-4389-a9c1-c115526eb52e) in Microsoft Teams (tied to your PSU AccessID)
2. Ensure that you have all of the proper software.  
    a. `R` (version 3.5.* minimum, version 4.0.0 preferred) [[Download](https://cloud.r-project.org/)] 
    b. RStudio Desktop(most current version preferred) [[Download](https://rstudio.com/products/rstudio/download/#download)]
3. Additional Software that we recommend  
    a. GitHub Desktop [[Download](https://desktop.github.com/)]
4. `R` Packages--here are some basic packages that everyone will need; be sure to install their dependencies too:

    ```{r, eval = FALSE}
    install.packages(c("devtools", "ggplot2", "shiny", "shinyBS", "tidyverse"))
    ```

5. A copy of the Sample App ([see below](#sample-app))

## `boastUtils` Package
Bob created the [boastUtils](https://github.com/EducationShinyAppTeam/boastUtils) package to automate much of the design and development process. This will not only reduce the amount of work you'll need to do, it'll also make apps more consistent. 
 
__Starting Summer 2020, you will be required to you make use of this tool.__

Please check out the package's [page](https://github.com/EducationShinyAppTeam/boastUtils) for instructions on installing and usage.

```{r, eval = FALSE}
library(devtools)
devtools::install_github("EducationShinyAppTeam/boastUtils")
```

## Sample App  
Bob and Neil have created a Sample App repository that you can use as a template for your own apps. To get started, clone the [Sample\_App](https://github.com/EducationShinyAppTeam/Sample_App) template repo found on GitHub. This will provide you with a skeleton for organizing your files as well as your code.

There are several methods you can use:

- [Direct Download (Basic)](#direct-download)
- [GitHub Desktop (Recommended)](#github-desktop)
- [Command Line (Advanced)](#command-line)

### Direct Download
You can download this repository directly by visiting:

> <https://github.com/EducationShinyAppTeam/Sample_App/archive/master.zip>

### GitHub Desktop
If you are using [GitHub Desktop](https://desktop.github.com/) and have [linked your account](https://help.github.com/en/desktop/getting-started-with-github-desktop/authenticating-to-github) that has access to the [EducationShinyAppTeam](https://github.com/EducationShinyAppTeam) repository, you can do the following from inside GitHub Desktop:  

1. Bring up the Clone Repository Menu (File -> Clone Repository...)  
2. Enter Sample\_App in the search bar and select the option that says Sample\_App (not sampleapp)
3. Click the Choose... button for the local path (this is where you want to the clone to live on your computer)
4. Click the Clone button.

**See also:**

- [GitHub Desktop Help](https://help.github.com/en/desktop)


### Command Line
Enter the following command in your terminal:

```{bash, eval = FALSE}
git clone git@github.com:EducationShinyAppTeam/Sample_App.git
```

**See also:**

- [Git Handbook](https://guides.github.com/introduction/git-handbook/)
- [Resources to learn Git](https://try.github.io/)
