# Documentation {#documentation}

These apps are the product of your hard work and are part of your academic record. Thus, you need to adhere to [Penn State's Academic Integrity Policy](https://undergrad.psu.edu/aappm/G-9-academic-integrity.html). This is especially important as we are making the apps available through a Creative Commons Attribution-ShareAlike 4.0 International license ([CC BY-SA 4.0](https://creativecommons.org/licenses/by-sa/4.0/)). If you have used code, pictures, data, or other materials from outside of the BOAST team, you __MUST__ give proper credit. These references will then be included on the App's References Tab.

## References
All apps will need a References Tab. This is where you'll place all references for your App, including R packages, borrowed code, data sources, images, etc. This in addition to the Acknowledgments.

__NOTE:__ listing something in the Acknowledgments DOES NOT waive this requirement.

You may use any citation style you wish, but be consistent. Recommended citation styles include APA and AmStat. Here is a starting code block for you to use:

```{r refTab1, echo=TRUE, eval=FALSE}
#[omitted code]
tabItem(
  tabName = "refs",
  withMathJax(),
  h2("References"),
  p(class = "hangingindent",
    "reference 1-alphabetically"),
  p(class = "hangingindent",
    "reference 2-alphabetically"),
  # Repeat as needed
)
```

Notice the use of `class = "hangingindent"`. You must include this with each reference as this will ensure the proper styling of your references.

If you need assistance with this section, please talk to Neil.

## Plagiarism

__You MAY NOT use blocks of code you've found online without giving proper attribution.__

There is a difference between looking at example code online to see how to do something and copying that code directly. The former is permissible, the later is plagiarism.

+ If you want to use someone else's code “as is” (without any changes), you should reach out to the author for permission first.
+ If you use someone else's code and make modifications, you need to give credit to where you got the code, and potentially ask for permission.

You will need to place citations in *__two__* places: in the References Tab and in your code. You might want to also consider adding an acknowledgement to the Overview tab.

### Reference Tab

Use the following format:

> Author. (Date). Title of program/source code (Version number, if applicable). [type of code]. Retrieved from \< URL \>.

For example,

> Hatfield, N. J. (2017). First day activity (v1). [Netlogo]. Retrieved from <https://neilhatfield.github.io/statApps/Day1Activity.html>.

### In Code

Use the following format in your code to cite where you got the code from.

```{r citeCode, echo=TRUE, eval=FALSE}
#-------------------------------------------------------------------------------
#  Title: <title of program/source code>
#  Author: <author(s) names>
#  Date: <date>
#  Code version: <code version>
#  Availability: <where it's located>
#-------------------------------------------------------------------------------
# [borrowed code then follows]
# ...
# [last line of borrowed code]      
#End of <author>'s code----------------------------------------------------------
```

## `R` Packages
If you made use of any packages in `R`, then you will need to add these to the Reference tab. Fortunately, there is a built-in tool that will help you: the `citation` function. In R (RStudio) simply type `citation("packageName")` and you'll get the appropriate citation information for the package you used. For example, `citation("shinydashboard")` and `citation("plyr")` will give the information needed for the following citations:

> Chang, W. and Borges Ribeio, B. (2018). shinydashboard: Create dashboards with ‘Shiny'. (v0.7.1) [R Package]. Available from <https://CRAN.R-project.org/package=shinydashboard>

> Wickham, H. (2011). The Split-apply-combine strategy for data analysis. Journal of Statistical Software, 40(1). pp. 1-29. Available at <http://www.jstatsoft.org/v40/i01/>.

Notice, that the format of the R package will depend on whether there is an article published for the package. The `shinydashboard` package is not associated with an article while the `plyr` package is associated with Wickham's article.


## Graphics
Pictures, drawings, photographs, images, etc. are typically copyrighted. When you're selecting images, make sure that the images are Open Source/Copyright Free/Royalty Free/Public Domain. Additionally, include a reference to where the pictures came from in the Overview Page. The basic format to use is:

> LastName, First Initial. (Year). Title of artwork. [Format]. Retrieved from \< URL \>.

## Data
If you are using any data files, you need to attribute where those files are coming from in the References tab. You might also want to add an acknowledgement on the Overview tab. A suggested format to use is:

>Author/Rightsholder. (Year). Title of data set (Version number) [Description of form]. Location: Name of producer.  

> Author/Rightsholder. (Year). Title of data set (Version number) [Description of form]. Retrieved from <http://www.url.com>

If you (or someone else) had to sign some type of agreement to access the data, we must examine the agreement before you make your App publicly accessible. Just because you got access to the data does not mean you have the right to share the data.

