# Static Images {#staticImages}

Design Style is the second side of the Visual Appearance of every app. Here, we deal with visual aspects that go beyond the layout of your App. This includes issues of branding, color, how you make text look (font, size, emphasis), and graphics (plots, tables, and images).

## PSU Branding {#logob}
Given that we are all associated with Pennsylvania State University, we need to include the Penn State logo in each App. Rather than sticking the logo at the top of the Overview page, we are going to place the logo at the bottom of the sidebar. This has the benefit of having the logo appear throughout the entire App AND making the logo be as unobtrusive as possible.

In your UI section of `app.R` (or the `ui.R` file), at the end of the `dashboardSidebar()` section, you will need to include the following:

```{r logoCodeb, echo=TRUE, eval=FALSE}
tags$div(class = "sidebar-logo",
         boastUtils::psu_eberly_logo("reversed"))
```

Here's how this code would look in context:

```{r psuLogob, echo=TRUE, eval=FALSE}
dashboardPage(
  skin = "blue",
  dashboardHeader(
    title = "title",
    # [omitted code]
  ),
  dashboardSidebar(
    width = 300,
    sidebarMenu(
      id = "tabs",
      # [omitted code],
      tags$div(
        class = "sidebar-logo",
        boastUtils::psu_eberly_logo("reversed")
        )
      ),
    dashboardBody(
    tags$head(
      tags$link(rel = "stylesheet", type = "text/css",
        href = "https://educationshinyappteam.github.io/Style_Guide/theme/boast.css"),
      tabItems(
       # [omitted code]
      )  )  )  )  )
```

This will ensure that the Penn State logo gets properly used.


## Colors

Your App needs to have a consistent color scheme throughout. The color scheme should be checked against colorblindness to meet [WCAG 2.1](https://www.w3.org/WAI/WCAG21/quickref/) Level AA. You can do so at the [Coloring for Colorblindness](https://davidmathlogic.com/colorblind/#%23000000-%23E69F00-%2356B4E9-%23009E73-%23F0E442-%230072B2-%23D55E00-%23CC79A7) website.  If you are following this Style Guide (as you should be) then the vast majority of this section will be automatically handled for you.

See [Section \@ref(colorUI)](#colorUI) and [Section \@ref(colorPlots)](#colorPlots) for additional information on the use of colors.

## Text Styling {#styleTextb}

Text styling refers the non-content aspects of the text on the page, such as the use of italics, boldface, alignment, as well as font size and color.

You should let the centralized CSS file do the heavy lifting for text styling. (Again, using `boastApp` will help you.) However, for this to work properly, you will need to tag content appropriately. (See the section on HTML, [Section \@ref(html)](#html).)

If you run into a situation where some element needs additional styling, __talk to Neil or Bob for help__. You might have come across an element that needs to get added the central CSS file or a bug.

### Headings

Use the Heading Tags for the short fragments that define the structure of your App. If you find yourself enclosing a complete sentence in Heading tag, you ARE NOT using headings correctly. Notice how the headings in this Style Guide aren’t complete sentences; your App should mimic this. Full sentences appear as regular paragraph text (i.e., enclosed in `p()`) and not be a Heading.

### Paragraph Text

If you enclose text that gives instructions or other information to your App’s users in `p()` or `li()` (the later should be wrapped in either `tags$ol()` or `tags$ul()`), your App will understand how to style that text correctly. The central CSS file contains controls that set the base font size much larger than Shiny does natively as well as making text sizing dynamic. (This is important for making our apps mobile device friendly.) Again, using `boastApp` makes this process easier.

If you want to make a certain word or phrase italic, you will need to wrap that text in `tags$em()`. Similarly, if you want do the same with boldface, you’ll use `tags$strong()`.

For example, this code:

```{r paraStyle1b, echo=TRUE, eval=FALSE}
p(
  "When dealing with the ",
  tags$em("t"),
  "-distribution, we only have one parameter, the ",
  tags$strong("degrees of freedom"),
  "that we need to input."
)
```

Becomes:  

> When dealing with the *t*-distribution, we only have one parameter, the __degrees of freedom__ that we need to input.

Use italics (emphasis), and boldface (strong) sparingly.

### Mathematics

For the most part, any mathematics you need displayed should be done using [MathJax](https://www.mathjax.org/). Default to using inline typesetting with the `\\(` and `\\)` delimiters. If you need to use display style, you can use `\\[` and `\\]`. For the vast majority of mathematics, you’ll wrap both inline and display style mathematics inside of a paragraph environment (`p()`).

If you’re writing mathematics directly in your app, remember you’ll need to escape the LaTeX commands by putting an extra backslash (\\) in front; e.g., `\frac{3}{4}` would need to be `\\frac{3}{4}`.

If you’re reading in mathematical text from an external CSV file, you do not need the extra backslash in the CSV file.

If you need assistance in figuring out how to type up mathematics, please talk to Neil, Matt, or Dennis.

__Note:__ Double dollar sign delimiters are generally not recommended for displaying math as they can lead to unintended results. See: [Writing Mathematics for MathJax](https://docs.mathjax.org/en/latest/basic/mathematics.html).

### [Game] Question Text

The text used as a question in a game should NOT be wrapped in a Heading tag; wrap the text in a paragraph tag.

### Label Text (Buttons, Sliders, Other Inputs and Alerts)

By using the central CSS file, any text you included in/on buttons, dropdown menus, sliders, radio buttons, choices, and other inputs as well as alert messages and popups/rollovers, will automatically be styled correctly.

Do not use heading tags, the paragraph tag, italics/emphasis, or boldface/strong with input labels. Input labels should be written in sentence case (i.e., capitalize only the first word and any proper nouns).

You may use these tags with popups/rollovers.

### Feedback and Hint Text

Again, let the central CSS file handle the styling of this type of text.

### Text in `R` Plots

Unfortunately, any text in `R` plots does not get controlled by CSS. This means that you’ll have to play around with the settings. Using the `ggplot2` package to make your plots (or other packages based upon the `grid` framework like `lattice`) will allow you to use the `theme` aspect to control text in your App.

Here is an example for how to do this:

```{r plotText1b, echo=TRUE, eval=FALSE}
# Create a ggplot2 object
g1 <- ggplot2::ggplot(data=df, aes(x=x, y=y, color=grp)) 
# Add your layers, for example
g1 + ggplot2::geom_point()
# Use theme to control text size
g1 + ggplot2::theme(
  plot.caption = element_text(size = 18),
  text = element_text(size = 18)
  )
```

You will need to play around with the settings to find the appropriate value; text size 18 appears to work out well in many cases.

__Note:__ The text in your plot might not behave well for dynamic re-sizing on different mobile devices.

### Text Case

When writing text, you should use the appropriate case (i.e., which words you capitalize):

+ Headings: use Title Case
+ Paragraphs/Lists: use Sentence Case
+ Input Labels: use Sentence Case
+ Figure Captions: use Title Case
+ Table Captions: use Title Case
+ Graphs (see [Section \@ref(graphLab)](#graphLab))
    - Titles: use Title Case
    - Axes: Use Sentence Case
    - Legends: Use Sentence Case
+ Default: Sentence Case

## Graphics {#graphicsb}

One of the most powerful tools we have in Statistics and Data Science is graphics. This includes images/pictures, graphs/plots, and tables. You will want to make sure that all graphical elements are appropriately sized in the Body. If there is text in a static image/picture, you'll need to make sure that the text is legible on a variety of screen sizes.

We've already discussed both issues of color and text size in plots. For additional considerations, please refer to the following readings (ordered from most important to least):

  - [Tufte-Fundamental Principles of Analytical Design](https://www.dropbox.com/s/hb52991v09p8q91/Tufte%20-%202006%20-%20The%20Fundamental%20principles%20of%20analytical%20design.pdf?dl=0)
  - [Tufte-Chartjunk](https://www.dropbox.com/s/z8yrf4eqph6c2h4/Tufte%20-%202001%20-%20Chartjunk%20Vibrations%2C%20grids%2C%20and%20ducks.pdf?dl=0)  
  - [Kosslyn-Looking with the Eye and Mind](https://www.dropbox.com/s/62uegsribwdjtze/Kosslyn%20-%202006%20-%20Looking%20with%20the%20eye%20and%20mind.pdf?dl=0)

Remember, we always want to be modeling excellent graphing behaviors.

> All photographs can be fortified with words. --Dorothea Lange

> A picture is worth a thousand words...but which ones? --Unknown

Both of these quotations highlight that you need to include some text with your plots to help the user construct their understanding of what you're trying to show them.

### Titles and Labels {#graphLabb}
Graph and Table titles should follow Title Case. Capitalize each word unless the word is "small" (e.g., of, an, etc.). The [Title Case website](https://titlecase.com/) can help you if you aren't sure. (This website also does other types of case such as camel case.)

### Axes and Scales
`R`'s default axes are terrible. They often do not fully cover the data and the have gaps between the axes. All this impedes the user's construction of meaning. Thus, you'll want to take control and stipulate the axes and scales to optimize what users get out of the plot. If you are providing multiple plots that the user is supposed to compare, make sure that they all use the same scaling and axes.

To force `ggplot2` to place (0, 0) in the lower-left corner and to control the scales, you will need to include the following:

```{r axesControl1b, echo=TRUE, eval=FALSE}
# Create the ggplot2 object
g1 <- ggplot2::ggplot(...)
# Add your layer
g1 + ggplot2::geom_point()
# Control axes and scale
## Multiplicative Scaling of the Horizontal (x) Axis
## Additive Scaling of the Vertical (y) Axis
g1 + ggplot2::scale_x_continuous(expand = expansion(mult = c(1,2), add = 0)) + 
  scale_y_continuous(expand = expansion(mult = 0, add = c(0,0.05))) 
```

### Color and Plots in `R` {#colorPlotsb}

In `R` you can set color theme which you use in `ggplot2`. Here are two custom color palettes that you can use in your App. Additionally, the package `viridis` provides several additional color palettes which are improvements upon the default color scheme.

```{r customPalettesb, echo=TRUE, eval=FALSE}
# boastPalette is based on the Wong color blind set found at the above website.
boastPalette <- c("#0072B2","#D55E00","#009E73","#CE77A8",
    "#000000","#E69F00","#999999","#56B4E9","#CC79A7")

# psuPalette is based on Penn State's three official color palettes
# and checked at the above webite.
psuPalette <- c("#1E407C","#BC204B","#3EA39E","#E98300",
    "#999999","#AC8DCE","#F2665E","#99CC00")

# Both palettes get used in the order of what is listed.
```

```{r boastPaletteb, fig.cap="The Boast Palette", fig.align='center', fig.width=7, fig.height=1.25, echo=FALSE}
boastPalette <- c("#0072B2","#D55E00","#009E73","#CE77A8",
    "#000000","#E69F00","#999999","#56B4E9","#CC79A7")

ggplot2::ggplot() +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = -8, xmax = -6,
      ymin = 0, ymax = 1),
    fill = boastPalette[1]
  ) +
  ggplot2::annotate(geom = "text", x = -7, y = 0.5,
                    label = boastPalette[1],
                    color = "white") +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = -6, xmax = -4,
      ymin = 0, ymax = 1),
    fill = boastPalette[2]
  ) +
  ggplot2::annotate(geom = "text", x = -5, y = 0.5,
                    label = boastPalette[2],
                    color = "white") +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = -4, xmax = -2,
      ymin = 0, ymax = 1),
    fill = boastPalette[3]
  ) +
  ggplot2::annotate(geom = "text", x = -3, y = 0.5,
                    label = boastPalette[3],
                    color = "white") +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = -2, xmax = -0,
      ymin = 0, ymax = 1),
    fill = boastPalette[4]
  ) +
  ggplot2::annotate(geom = "text", x = -1, y = 0.5,
                    label = boastPalette[4],
                    color = "white") +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = 0, xmax = 2,
      ymin = 0, ymax = 1),
    fill = boastPalette[5]
  ) +
  ggplot2::annotate(geom = "text", x = 1, y = 0.5,
                    label = boastPalette[5],
                    color = "white") +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = 2, xmax = 4,
      ymin = 0, ymax = 1),
    fill = boastPalette[6]
  ) +
  ggplot2::annotate(geom = "text", x = 3, y = 0.5,
                    label = boastPalette[6],
                    color = "black") +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = 4, xmax = 6,
      ymin = 0, ymax = 1),
    fill = boastPalette[7]
  ) +
  ggplot2::annotate(geom = "text", x = 5, y = 0.5,
                    label = boastPalette[7],
                    color = "black") +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = 6, xmax = 8,
      ymin = 0, ymax = 1),
    fill = boastPalette[8]
  ) +
  ggplot2::annotate(geom = "text", x = 7, y = 0.5,
                    label = boastPalette[8],
                    color = "black") +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = 8, xmax = 10,
      ymin = 0, ymax = 1),
    fill = boastPalette[9]
  ) +
  ggplot2::annotate(geom = "text", x = 9, y = 0.5,
                    label = boastPalette[9],
                    color = "black") +
  ggplot2::theme_void()

```


```{r psuPaletteb, fig.cap="The PSU Palette", fig.align='center', fig.width=7, fig.height=1.25, echo=FALSE}
psuPalette <- c("#1E407C","#BC204B","#3EA39E","#E98300",
    "#999999","#AC8DCE","#F2665E","#99CC00")

ggplot2::ggplot() +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = -8, xmax = -6,
      ymin = 0, ymax = 1),
    fill = psuPalette[1]
  ) +
  ggplot2::annotate(geom = "text", x = -7, y = 0.5,
                    label = psuPalette[1],
                    color = "white") +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = -6, xmax = -4,
      ymin = 0, ymax = 1),
    fill = psuPalette[2]
  ) +
  ggplot2::annotate(geom = "text", x = -5, y = 0.5,
                    label = psuPalette[2],
                    color = "white") +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = -4, xmax = -2,
      ymin = 0, ymax = 1),
    fill = psuPalette[3]
  ) +
  ggplot2::annotate(geom = "text", x = -3, y = 0.5,
                    label = psuPalette[3],
                    color = "white") +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = -2, xmax = -0,
      ymin = 0, ymax = 1),
    fill = psuPalette[4]
  ) +
  ggplot2::annotate(geom = "text", x = -1, y = 0.5,
                    label = psuPalette[4],
                    color = "white") +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = 0, xmax = 2,
      ymin = 0, ymax = 1),
    fill = psuPalette[5]
  ) +
  ggplot2::annotate(geom = "text", x = 1, y = 0.5,
                    label = psuPalette[5],
                    color = "white") +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = 2, xmax = 4,
      ymin = 0, ymax = 1),
    fill = psuPalette[6]
  ) +
  ggplot2::annotate(geom = "text", x = 3, y = 0.5,
                    label = psuPalette[6],
                    color = "black") +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = 4, xmax = 6,
      ymin = 0, ymax = 1),
    fill = psuPalette[7]
  ) +
  ggplot2::annotate(geom = "text", x = 5, y = 0.5,
                    label = psuPalette[7],
                    color = "black") +
  ggplot2::geom_rect(
    mapping = ggplot2::aes(
      xmin = 6, xmax = 8,
      ymin = 0, ymax = 1),
    fill = psuPalette[8]
  ) +
  ggplot2::annotate(geom = "text", x = 7, y = 0.5,
                    label = psuPalette[8],
                    color = "black") +
  ggplot2::theme_void()
```

To use these palettes (or ones from `viridis`) with a `ggplot2` object, you'll need to do the following

```{r examplePaletteb, echo=TRUE, eval=FALSE}
# You will need to first add whichever palette line from above to your code
boastPalette <- c("#0072B2","#D55E00","#009E73","#CE77A8",
    "#000000","#E69F00","#999999","#56B4E9","#CC79A7")

# Create ggplot2 object
g1 <- ggplot2::ggplot(data = df, 
                aes(x = x, y = y, color = grp, fill = grp))
# Add your layers
g1 + ggplot2::geom_points()
# Tell R to use your chosen palette
g1 + ggplot2::scale_color_manual(values=boastPalette)  # If you use "color" in aes
g1 + ggplot2::scale_fill_manual(values=boastPalette)  # If you use "fill" in aes
```

If you have more groups than eight/nine colors listed in the two palettes, consider reworking your examples as you could overwhelm the user with too many colors. (This also applies to using different shapes to plot points.)

With an eye towards accessibility, try not to use only color to denote a particular piece of information. Rather you might want to use color and shape.

### Tables
Data tables can pose a challenge for individuals to comprehend. Just as a wall of text isn't conducive to helping a person understand what's going on, neither is a wall of data values. Thus, we need to be extreme judicious (picky) about incorporating data tables into any of our apps.

In web development there are two main types of tables: layout tables and data tables. 

+ Layout tables help control where different elements appear on the page. 
+ We need an additional distinction for data tables:
    - Summary Data Tables are tables that have summary information; typified by two-way tables (a.k.a. contingency tables or crosstabs) but might also include other things such as values of descriptive statistics stratified by groups. 
    - Data Sets are an entire data object, presented in tabular format

__Layout Tables should never be used in a BOAST App.__

Data Sets should be displayed __as sparingly as possible__. In order to include a Data Set display, you will need to have identified an explicit learning goal/objective that necessitates the user digging through a data frame. If you can't identify such a learning goal, you should NOT include a data frame.

If the goal is to allow the user to look through the data set OR to have access to the data, then __give a link__ to either the original source of the data (__preferred__) or for them to download the file.

Summary Data Tables can be used more often and can enrich the user's experience with your app. However, these must still be constructed in an appropriate manner.

Neither Data Sets nor Summary Data Tables should be inserted into your App as a picture. This is an big Accessibility violation. Use the directions below to create the appropriate type of data table.

#### Displaying Data Tables
Your first step is to create a data frame object in your `R` code. 

+ If you are displaying a data set (rare), then you will either need to read in the data or call that data frame. For this example, we'll be using the `mtcars` data frame that is part of `R`. 
+ If you are making a Summary Data Table, you'll need to either use `R` to calculate the values and store in a data frame or create a data frame yourself.

In either case, be sure you identify what columns you're going to use. If your original data file has 50 columns, but your App only makes use of 5, drop the other 45. Only display the columns that you actually use.

Your next step is to decide on where to put this display (e.g., inside an Exploration Tab or as a separate page). This will help you identify where in your App's UI section you need to put the appropriate code. 

To ensure that your data table is accessible and responsive (i.e., mobile friendly), you will need to use the `DT` package.
```{r installDTb, echo=TRUE, eval=FALSE}
install.packages("DT")
# Be sure to include this in your library call
library(DT)
```

In your UI section, you'll need to use the following code, placed in the appropriate area:
```{r dtUICode1b, echo=TRUE, eval=FALSE}
# [code omitted]
DT::DTOutput(outputId = "mtCars")
# [code omitted]
```

Then, in your Server section, you'll need to use the following code:
```{r dtServerCode1b, echo=TRUE, eval=FALSE}
# [code omitted]
# Prepare your data set with only the columns needed
carData <- mtcars[,c("mpg", "cyl", "hp", "gear", "wt")]

## Use Short but Meaningful Column Names
names(carData) <- c("MPG", "# of Cylinders", "Horsepower", "# of Gears", "Weight")

# Create the output data table
# Be sure to use the same name as you did in the UI
output$mtCars <- DT::renderDT(
  expr = carData,
  caption = "Motor Trend US Data, 1973-1974 Models", # Add a caption to your table
  style = "bootstrap4", # You must use this style
  rownames = TRUE,
  options = list( # You must use these options
    responsive = TRUE, # allows the data table to be mobile friendly
    scrollX = TRUE, # allows the user to scroll through a wide table
    columnDefs = list(  # These will set alignment of data values
      # Notice the use of ncol on your data frame; leave the 1 as is.
      list(className = 'dt-center', targets = 1:ncol(carData))
    )
  )
)
# [code omitted]
```

If you are making a Summary Data Table, you will need to follow the same process. If your data frame does not have row names, but instead a column with values acting as row names, you may replace the `rownames = TRUE` with `rownames = FALSE`; there should not be a column of sequential numbers on the left.

Column names __MUST__ be simple and *meaningful* to the user. To this end, you should rename any columns that might have poor choices for names, just as we have done with the `mtcars` data. This includes using Greek characters in isolation. You should not have any columns labeled \(\mu\) or \(\sigma\). Rather you need to use English words.

*Note: getting mathematical expressions to render properly in graphical environments in `R` is not as easy as in the paragraphs or headers of an app. Only certain graphing packages support limited mathematical expressions. The same is true for table generation packages.* 

Again, try to use tables as infrequently as possible. Poorly constructed tables can create accessibility issues causing screen readers to poorly communicate tables to your users. If you run into problems and/or have questions, __talk to Neil and Bob__. 

#### Additional Table Examples

We're including some additional Summary Data Table examples. For these examples, I'm going to make use of the `palmerpenguins` package of data sets.

##### Summary Data Table of Descriptive Statistics

```{r dtSummary1Setupb, echo=TRUE, eval=TRUE}
library(palmerpenguins)
library(psych)
library(DT)
library(tibble)

penStats <- psych::describeBy(
  x = penguins$body_mass_g,
  group = penguins$species,
  mat = TRUE, # Formats output appropriate for DT
  digits = 3 # sets the number of digits retained
)

# Picking which columns to keep
penStats <- penStats[, c("group1", "n", "mean", "sd", "median", "mad", "min", "max", "skew",
                         "kurtosis")]
# Make the group1 column the row names
penStats <- tibble::remove_rownames(penStats)
penStats <- tibble::column_to_rownames(penStats,
                          var = "group1")
# Improve column names
names(penStats) <- c("Count", "SAM (g/penguin)", "SASD (g)", "Median (g)", "MAD (g)", "Min (g)",
                     "Max (g)", "Sample Skewness (g^3)", "Sample Excess Kurtosis (g^4)")
```

```{r dtSummary1Shinyb, echo=TRUE, eval=FALSE}
# Make the Table
output$penguinSummary <- DT::renderDT(
  expr = penStats,
  caption = "Descriptive Stats for Palmer Penguins", 
  style = "bootstrap4", 
  rownames = TRUE,
  autoHideNavigation = TRUE,
  options = list( 
    responsive = TRUE, 
    scrollX = TRUE,
    paging = FALSE, # Set to False for small tables
     columnDefs = list(
       list(className = 'dt-center',
            targets = 1:ncol(penStats))
    )
  )
)
```

```{r dtSummary1Targetb, echo=FALSE, eval=FALSE}
DT::datatable(
  data = penStats,
  caption = "Descriptive Stats for Palmer Penguins", 
  style = "bootstrap4", 
  rownames = TRUE,
  autoHideNavigation = TRUE,
  options = list( 
    responsive = TRUE, 
    scrollX = TRUE,
    paging = FALSE,
     columnDefs = list(
       list(className = 'dt-center',
            targets = 1:ncol(penStats))
    )
  )
)

```

##### Summary Data Table for Output Table

While this example is for an ANOVA table, you can build from this for other output tables. If you store the output of any call as an object, you can then use the structure function, `str` to investigate the output. Ultimately, you need something that is either a matrix or a data frame. 

```{r dtSummary2Setupb, echo=TRUE, eval=TRUE}
library(palmerpenguins)
library(psych)
library(DT)
library(tibble)
library(rstatix)

# This bad practice but I'm going to pretend that all assumptions are met
penModel <- aov(body_mass_g ~ species*sex, data = penguins)

anovaPen <- round(anova(penModel), 3)
# Rounding to truncate decimals

```

```{r dtSummary2Shinyb, echo=TRUE, eval=FALSE}
# Make the Table
output$penguinAnova <- DT::renderDT(
  expr = anovaPen,
  caption = "(Classical) ANOVA Table for Palmer Pengins", 
  style = "bootstrap4", 
  rownames = TRUE,
  options = list( 
    responsive = TRUE, 
    scrollX = TRUE,
    paging = FALSE, # Set to False for small tables
     columnDefs = list(
       list(className = 'dt-center',
            targets = 1:ncol(anovaPen))
    )
  )
)
```

```{r dtSummary2Targetb, echo=FALSE, eval=FALSE}
DT::datatable(
  data = anovaPen,
  caption = "(Classical) ANOVA Table for Palmer Penguins", 
  style = "bootstrap4", 
  rownames = TRUE,
  options = list( 
    responsive = TRUE, 
    scrollX = TRUE,
    paging = FALSE,
     columnDefs = list(
       list(className = 'dt-center',
            targets = 1:ncol(anovaPen))
    )
  )
)
```


### Static Images
Static image refers to any image you're using in your App which is not produced by `R`. These are usually PNG or JPG/JPEG files which you end up calling in the UI portion of your code.

Within your App's folder/directory, there needs to be a sub-folder/directory called `www`. This is the place where you'll need to place ALL static image files.

#### Adding an Image

To include the image in your App, you'll need to make use of the image tag, `img`. When you run your App, Shiny automatically knows to check the `www` folder any time the `img` tag gets called. Here is an example supposing that the check mark image for correct answers is in the app's `www` folder:

```{r imageEx1b, echo=TRUE, eval=FALSE}
#[code omitted]
div(align = "right",
    img(src = "check.PNG",
        alt = "Success, you are correct",
        width = 25, #these are in pixels
        height = 25,
        ))
#[code omitted]

```

You'll notice that we've wrapped the `img` call in a `div` call. The `div` call allows us to specify that we want the image to be right aligned; you could also do left or center. You can replace the `div` call with the paragraph environment and include text on either side, effectively making your image part of the text.

```{r imageEx2b, echo=TRUE, eval=FALSE}
#[code omitted]
p("Check your answer here -->",
  img(src = "check.PNG",
      alt = "Success, you are correct",
      width = 25, height = 25),
  "<-- Check your answer here"),
#[code omitted]

```

#### Sizing and Positioning Your Image
All image files have a native size that is part of that file. For instance, the check mark image is 270 x 250 pixels. However, we overrode that that sizing with the `width` and `height` arguments. How did we decide on 25 x 25? Honestly, through guessing and checking. You'll need to think about how you're using the image and let that guide your decision making. There is no one size fits all solution. While finding an optimal size and position for your image can take some time, seeing bad settings is pretty obvious. Feel free to reach out to Neil and Bob for assistance.

### Alt Text {#altTextb}
Any graphical element you include in your App __MUST__ have an alternative (assistive) text description ("alt text"). This provides a short description of what is in the image or plot for users who are visual impaired. (Tables, when properly formatted will handle this automatically.)


Here are several resources worth checking out:

+ [WebAIM Alternative Text Guide](https://webaim.org/techniques/alttext/#basics)
+ [Penn State's Image ALT Text Page](https://accessibility.psu.edu/images/alttext/)
+ [W3C's ALT Text Decision Tree](https://www.w3.org/WAI/tutorials/images/decision-tree/)

#### Adding Alt Text to Static Images {#altStaticb}

In the prior section on static images, you saw exactly how to set the alt text; here is a generic example:

```{r altText1b, echo=TRUE, eval=FALSE}
#[code omitted]
img(src = "yourImage.PNG",
    alt = "Short description of what's in the pic",
    width = 25, height = 25)
#[code omitted]

```

#### Adding Alt Text Graphs

At this point in time (6/10/2020), adding alt text to plots generated in `R` is not as easy as for static pictures. While there are some potential changes coming down the pipe, there is not a firm time line for the addition of appropriate arguments to the existing `render*` functions. Therefore we will need to put in some (hopefully) temporary measures.

In particular, we will use [Accessible Rich Internet Applications (ARIA)](https://developer.mozilla.org/en-US/docs/Web/Accessibility/ARIA) to assist us in writing some labels that will stand in place of formal alt text. To do this, you will need to make use of the following code:

```{r plotsAltTextb, echo=TRUE, eval=FALSE}
# [code omitted]
# In the UI section, in the appropriate tabItem
plotOutput(outputId = plotID) # Look for lines like this 
# Code for adding the aria label
tags$script(HTML(
  "$(document).ready(function() {
  document.getElementById('plotId').setAttribute('aria-label',
  `General description of the plot`)
  })"
))
# [code omitted]
```

Important things to note:

1. Place the `tags$script(HTML(...))` code right after each instance of `plotOutput`.
2. Copy the above code as formatted
3. Change the two (2) pieces for each particular plot
    a. Replace `plotId` (keep the single quotation marks in the code)
    b. Replace `General description of the plot` (keep the single quotation marks in the code)

